---
title: "Working with Likert-scale data"
description: |
  A short description of the post.
author:
  - name: Lefkios  Paikousis
    url: https://www.linkedin.com/in/lefkios/
date: 01-03-2021
output:
  distill::distill_article:
    self_contained: false
draft: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


### Libraries

```{r message=FALSE, warning=FALSE}

library(labelled)
library(tidyverse)

```


The data come from the __UCI Machine Learning Repository__ and the direct link for the data and description can be found [here](https://archive.ics.uci.edu/ml/datasets/wiki4he).

The data reference ^[Meseguer‐Artola, A., Aibar, E., Lladós, J., Minguillón, J. and Lerga, M., 2015. Factors that influence the teaching use of Wikipedia in higher education. Journal of the Association for Information Science and Technology, 67(5), pp.1224-1232.ISSN: 2330-1635. doi: 10.1002/asi.23488](Meseguer‐Artola et al., 2015)




```{r}
path <- "https://archive.ics.uci.edu/ml/machine-learning-databases/00334/wiki4HE.csv"

spps_path <-"https://www.sheffield.ac.uk/polopoly_fs/1.886067!/file/Video_SPSS.sav"

# Missing values in the data are represented as `?`s
dta <- read_delim(path, delim = ";", na = c("", NA, "?"))

glimpse(dta)
```


```{r}

labels_vector <- c(

# Perceived Usefulness
"PU1"  = "The use of Wikipedia makes it easier for students to develop new skills",
"PU2"  = "The use of Wikipedia improves students' learning",
"PU3"  = "Wikipedia is useful for teaching",

#Perceived Ease of Use
"PEU1" = "Wikipedia is user-friendly",
"PEU2" = "It is easy to find in Wikipedia the information you seek",
"PEU3" = "It is easy to add or edit information in Wikipedia",

# Quality
"Qu1" = "Articles in Wikipedia are reliable",
"Qu2" = "Articles in Wikipedia are updated",
"Qu3" = "Articles in Wikipedia are comprehensive",
"Qu4" = "In my area of expertise, Wikipedia has a lower quality than other educational resources",
"Qu5" = "I trust in the editing system of Wikipedia"
)

labels_list <- labels_vector %>% as.list()

```

```{r}

dta %>% count(PU1)


labelled::var_label(dta) <- labels_list


likert_levels <- c(
  "Strongly disagree"  = "1",
  "Disagree" = "2",
  "Neutral" = "3",
  "Agree" = "4",
  "Strongly agree"  = "5"
)

domain_levels <- c(
  
  "Arts & Humanities" = "1",
  "Sciences" = "2",
  "Health Sciences" = "3",
  "Engineering & Architecture" = "4",
  "Law & Politics" = "5"
  
)

        
  # Recode to "Agree".. "Disagree"

dta %>%
  mutate(
    across(all_of(names(labels_list)), ~{
      fct_recode(as.factor(.), !!!likert_levels)
    })
  ) %>% 
  
  # Recode Gender and Donain
  mutate(
    gender = case_when(
                       GENDER == 1 ~ "Female",
                       GENDER == 0 ~ "Male",
                       TRUE ~ NA_character_
                       ),
    domain = fct_recode(as.factor(DOMAIN), !!!domain_levels)
            
  ) 


```


