---
title: "CSS tips for Shiny"
description: |
  A collection of tips/ tricks for using CSS to tweak the UI of a shiny app. These are solutions
  I mostly look for when building shiny apps and want to improve the UI
author:
  - name: Lefkios  Paikousis
    url: https://www.linkedin.com/in/lefkios/
categories: 
  - CSS
  - Shiny
date: 2022-08-06
output:
  distill::distill_article:
    toc: true
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

Here I will outline a collection of tips for using CSS to tweak the UI of a shiny app.

I find myself looking on Google for the same things, over and over again, so I decided to create a post to outline all the tips and tricks or resources when I want to do something with CSS on my Shiny app.

I also hope that other people will benefit from this.

The code for all the examples can be found on the [posts' github page](https://github.com/lefkiospaikousis/distill_webpage/tree/master/_posts/2022-08-06-css-tips-for-shiny)

### Center a button (horizontaly)

OMG.. how many times I need to search for this

`actionButton("go", "Go", style = "display:block; margin:0 auto")`

This is the [inline style](https://shiny.rstudio.com/articles/css.html#:~:text=pros%20and%20cons.-,Inline%20CSS,-The%20quickest%20and).

The best is to use the [File-based](https://shiny.rstudio.com/articles/css.html#:~:text=file%2Dbased%20workflow) CSS

Let's say we add a class to the button called `.center` (or whatever class name you want)

    actionButton("go", "Go", class = ".center')

and on the CSS file, we add:

    .center{
      
      display:block; 
      margin: 0 auto;
    }

To find out how to center both horizontaly and vertically, [visit this](https://sebhastian.com/css-center-button/)

### Scroll bar in Modal Dialogue

CSS ans SO to the rescue. [See here](https://stackoverflow.com/a/25874635/8689932)

In the CSS file, we add:

    .modal-dialog { 
      overflow-y: initial;
    }

    .modal-body{
      max-height: 80vh;
      overflow-y: auto;
    }

Note that the classes names are __exactly__ these (`modal-dialog`, `modal-body`.   
when you type in the console:

`shiny::modalDialog()`

You can see that these are the classes

```{r echo=FALSE}
print(shiny::modalDialog())
```

__Questions:__

- How to I apply this to a single modal?
- How do I make the `max-height` property variable to the screen size? (since I assume this is for desktops)







